### 2019-11-13 analysis/output/include/clust_summ.tex
Made tikz path diagrams

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Fri, Nov 08, 2019 - 11:40:45
\begin{table}[!htbp] \centering 
  \caption{} 
  \label{clust_summ} 
\begin{tabular}{@{\extracolsep{5pt}} ccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
clust5 & c\_ind & sp\_rich\_raref & n\_stems\_ha & bchave \\ 
\hline \\[-1.8ex] 
C1 & Diplorhynchus condylocarpon, Combretum spp., Pseudolachnostylis maprouneifolia & 7+4.7 & 276+345.2 & 35.3+37.03 \\ 
C2 & Julbernardia paniculata, Isoberlinia angolensis, Albizia antunesiana & 8+4.3 & 274+245.9 & 46.8+35.05 \\ 
C3 & Burkea africana, Pterocarpus angolensis, Baikiaea plurijuga & 4+2.3 & 274+536.5 & 36.3+27.86 \\ 
C4 & Baikiaea plurijuga, Terminalia randii, Albizia amara & 10+5.9 & 447+386.3 & 47+45.41 \\ 
C5 & Colophospermum mopane, Pseudolachnostylis maprouneifolia & 4+2.9 & 322+390.1 & 43.8+34.45 \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}