### 2019-11-13 manuscript/include/path_diagram_full.tex

\begin{tikzpicture}[auto,scale=2, 
	observed/.style={rectangle,draw,thick,inner sep=5pt,minimum size=1cm, align=center},
	latent/.style={circle, draw, thick, inner sep=0pt, minimum size=2cm, align=center}, 
	path/.style={->, thick, >=stealth'},
	loading/.style={->, thick, dashed, >=stealth'}]

\tikzset{mystyle/.style={->,double=black}}
\node [latent] (m) {Moisture};
\node [latent] (d) [below = 1in of m] {Species\\Diversity}; 
\node [latent] (s) [below = 1in of d] {Soil}; 
\node [latent] (h) [right = 1in of d] {Struc.\\Diversity}; 
\node [observed] (b) [right = 1in of h] {AGB}; 
\node [observed] (i) [left = 1.5in of d] {Stem\\Density};

\coordinate[above= 0.7in of m] (momid);
\node [observed] (mp) [left = 0.8in of momid] {MAP};
\node [observed] (mpc) [left = 0.1in of momid] {MAP\\CoV};
\node [observed] (mt) [right = 0.1in of momid] {MAT};
\node [observed] (mtc) [right = 0.8in of momid] {MAT\\CoV};

\coordinate[below= 0.7in of s] (somid);
\node [observed] (sc) [left = 0.5in of somid] {SOC};
\node [observed] (ss) at (somid) {Sand};
\node [observed] (so) [right = 0.5in of somid] {CEC};

\coordinate[left= 0.3in of d] (domid);
\node [observed] (ds) [above = 0.4in of domid] {Sp.\\Rich.};
\node [observed] (de) [below = 0.42in of domid] {Shannon\\Equit.};

\coordinate[left= 0.25in of h] (homid); 
\node [observed] (hd) [above = 0.4in of homid] {DBH\\CoV};
\node [observed] (hh) [below = 0.4in of homid] {Height\\CoV};

\draw [path] (m) to node {\pcfmd{}} (d);
\draw [path] (s) to node[right] {\pcfsd{}} (d);
\draw [path] (d) to node[below] {\pcfdh{}} (h);
\draw [path] (h) to node {\pcfhb{}} (b);
\draw [path] (m) to node[above=0.1in, pos=0.5] {\pcfmi{}} (i);
\draw [path] (s) to node {\pcfsi{}} (i);
\draw [path] (d) to node {\pcfdi{}} (i);
\draw [path] (m) to node {\pcfmb{}} (b);
\draw [path] (s) to node[below=0.1in, pos=0.5] {\pcfsb{}} (b);

\draw [loading] (m) to node[left, pos=0.5] {\pcfmmp{}} (mp);
\draw [loading] (m) to node[left, pos=0.5] {\pcfmmpc{}} (mpc);
\draw [loading] (m) to node[right, pos=0.5] {\pcfmmt{}} (mt);
\draw [loading] (m) to node[right, pos=0.5] {\pcfmmtc{}} (mtc);

\draw [loading] (s) to node[left, pos=0.5] {\pcfssc{}} (sc);
\draw [loading] (s) to node[left, pos=0.5] {\pcfsss{}} (ss);
\draw [loading] (s) to node[right, pos=0.5] {\pcfsso{}} (so);

\draw [loading] (d) to node {\pcfdds{}} (ds);
\draw [loading] (d) to node[left, pos=0] {\pcfdde{}} (de);

\draw [loading] (h) to node[] {\pcfhhd{}} (hd);
\draw [loading] (h) to node {\pcfhhh{}} (hh);

\coordinate[below= 3in of i] (ibleft); 
\coordinate[below= 3in of b] (ibright); 
\draw [path] (i) -- (ibleft) to node[below] {\pcfib{}} (ibright) -- (b);

\coordinate[above= 1.3in of b] (dbright);
\coordinate[left= 2in of dbright] (dbleft);
\draw [path] (d) -- (dbleft) to node[above] {\pcfdb{}} (dbright) -- (b);

\end{tikzpicture}