### 2019-11-19 analysis/output/include/clust_summ.tex
Finished editing

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Tue, Nov 19, 2019 - 14:43:03
\begin{table}[!htbp] \centering 
  \caption{} 
  \label{clust_summ} 
\begin{tabular}{@{\extracolsep{5pt}} cccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
clust5 & c\_ind & n\_plots & sp\_rich\_raref & n\_stems\_ha & bchave \\ 
\hline \\[-1.8ex] 
C1 & Diplorhynchus condylocarpon, Combretum spp., Pseudolachnostylis maprouneifolia & 478 & 7(5) & 316(228.6) & 37.1(27.66) \\ 
C2 & Julbernardia paniculata, Isoberlinia angolensis, Albizia antunesiana & 617 & 8(5) & 291(191.7) & 43.8(40.28) \\ 
C3 & Burkea africana, Pterocarpus angolensis, Baikiaea plurijuga & 62 & 5(2) & 362(259.4) & 37(42.59) \\ 
C4 & Baikiaea plurijuga, Terminalia randii, Albizia amara & 219 & 11(9) & 495(516.8) & 51.7(50.74) \\ 
C5 & Colophospermum mopane, Pseudolachnostylis maprouneifolia & 65 & 5(5) & 424(226.8) & 50.9(32.7) \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}