### 2019-11-20 analysis/output/include/corr_ci_tab.tex

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Wed, Nov 20, 2019 - 15:22:31
\begin{table}[!htbp] \centering 
  \caption{} 
  \label{corr_ci_tab} 
\begin{tabular}{@{\extracolsep{5pt}} ccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
x\_var & y\_var & raw.r & raw.lower & raw.upper & n & p \\ 
\hline \\[-1.8ex] 
sand\_per\_std & ocdens\_std & $$-$0.500$ & $$-$0.540$ & $$-$0.470$ & $1766$ & $0$ \\ 
sand\_per\_std & cation\_ex\_cap\_std & $$-$0.560$ & $$-$0.600$ & $$-$0.530$ & $1766$ & $0$ \\ 
sand\_per\_std & total\_precip\_std & $$-$0.490$ & $$-$0.520$ & $$-$0.450$ & $1766$ & $0$ \\ 
sand\_per\_std & precip\_seasonality\_log\_std & $0.330$ & $0.280$ & $0.370$ & $1766$ & $0$ \\ 
sand\_per\_std & mean\_temp\_std & $0.290$ & $0.250$ & $0.330$ & $1766$ & $0$ \\ 
sand\_per\_std & temp\_seasonality\_log\_std & $0.220$ & $0.180$ & $0.270$ & $1766$ & $0$ \\ 
sand\_per\_std & sp\_rich\_raref\_log\_std & $$-$0.360$ & $$-$0.400$ & $$-$0.310$ & $1766$ & $0$ \\ 
sand\_per\_std & shannon\_equit\_log\_std & $$-$0.300$ & $$-$0.340$ & $$-$0.260$ & $1766$ & $0$ \\ 
sand\_per\_std & cov\_height\_std & $$-$0.120$ & $$-$0.170$ & $$-$0.060$ & $1326$ & $0$ \\ 
sand\_per\_std & cov\_dbh\_std & $$-$0.190$ & $$-$0.230$ & $$-$0.140$ & $1766$ & $0$ \\ 
sand\_per\_std & stems\_ha\_log\_std & $$-$0.270$ & $$-$0.310$ & $$-$0.230$ & $1766$ & $0$ \\ 
sand\_per\_std & bchave\_log & $$-$0.230$ & $$-$0.280$ & $$-$0.190$ & $1766$ & $0$ \\ 
ocdens\_std & cation\_ex\_cap\_std & $0.300$ & $0.250$ & $0.340$ & $1766$ & $0$ \\ 
ocdens\_std & total\_precip\_std & $0.440$ & $0.400$ & $0.480$ & $1766$ & $0$ \\ 
ocdens\_std & precip\_seasonality\_log\_std & $$-$0.310$ & $$-$0.350$ & $$-$0.270$ & $1766$ & $0$ \\ 
ocdens\_std & mean\_temp\_std & $$-$0.310$ & $$-$0.350$ & $$-$0.270$ & $1766$ & $0$ \\ 
ocdens\_std & temp\_seasonality\_log\_std & $$-$0.140$ & $$-$0.190$ & $$-$0.100$ & $1766$ & $0$ \\ 
ocdens\_std & sp\_rich\_raref\_log\_std & $0.260$ & $0.210$ & $0.300$ & $1766$ & $0$ \\ 
ocdens\_std & shannon\_equit\_log\_std & $0.220$ & $0.180$ & $0.270$ & $1766$ & $0$ \\ 
ocdens\_std & cov\_height\_std & $0.070$ & $0.020$ & $0.130$ & $1326$ & $0.006$ \\ 
ocdens\_std & cov\_dbh\_std & $0.140$ & $0.090$ & $0.190$ & $1766$ & $0$ \\ 
ocdens\_std & stems\_ha\_log\_std & $0.170$ & $0.120$ & $0.210$ & $1766$ & $0$ \\ 
ocdens\_std & bchave\_log & $0.250$ & $0.200$ & $0.290$ & $1766$ & $0$ \\ 
cation\_ex\_cap\_std & total\_precip\_std & $$-$0.010$ & $$-$0.060$ & $0.040$ & $1766$ & $0.657$ \\ 
cation\_ex\_cap\_std & precip\_seasonality\_log\_std & $$-$0.510$ & $$-$0.540$ & $$-$0.480$ & $1766$ & $0$ \\ 
cation\_ex\_cap\_std & mean\_temp\_std & $0.190$ & $0.150$ & $0.240$ & $1766$ & $0$ \\ 
cation\_ex\_cap\_std & temp\_seasonality\_log\_std & $0.030$ & $$-$0.020$ & $0.070$ & $1766$ & $0.260$ \\ 
cation\_ex\_cap\_std & sp\_rich\_raref\_log\_std & $$-$0.020$ & $$-$0.060$ & $0.030$ & $1766$ & $0.464$ \\ 
cation\_ex\_cap\_std & shannon\_equit\_log\_std & $0.220$ & $0.170$ & $0.260$ & $1766$ & $0$ \\ 
cation\_ex\_cap\_std & cov\_height\_std & $0.040$ & $$-$0.010$ & $0.090$ & $1326$ & $0.149$ \\ 
cation\_ex\_cap\_std & cov\_dbh\_std & $0.240$ & $0.190$ & $0.280$ & $1766$ & $0$ \\ 
cation\_ex\_cap\_std & stems\_ha\_log\_std & $0.200$ & $0.150$ & $0.240$ & $1766$ & $0$ \\ 
cation\_ex\_cap\_std & bchave\_log & $0.030$ & $$-$0.020$ & $0.080$ & $1766$ & $0.230$ \\ 
total\_precip\_std & precip\_seasonality\_log\_std & $$-$0.120$ & $$-$0.160$ & $$-$0.070$ & $1766$ & $0$ \\ 
total\_precip\_std & mean\_temp\_std & $$-$0.120$ & $$-$0.170$ & $$-$0.070$ & $1766$ & $0$ \\ 
total\_precip\_std & temp\_seasonality\_log\_std & $$-$0.690$ & $$-$0.720$ & $$-$0.670$ & $1766$ & $0$ \\ 
total\_precip\_std & sp\_rich\_raref\_log\_std & $0.430$ & $0.390$ & $0.470$ & $1766$ & $0$ \\ 
total\_precip\_std & shannon\_equit\_log\_std & $0.100$ & $0.060$ & $0.150$ & $1766$ & $0$ \\ 
total\_precip\_std & cov\_height\_std & $0.250$ & $0.200$ & $0.300$ & $1326$ & $0$ \\ 
total\_precip\_std & cov\_dbh\_std & $0.150$ & $0.100$ & $0.190$ & $1766$ & $0$ \\ 
total\_precip\_std & stems\_ha\_log\_std & $0.070$ & $0.030$ & $0.120$ & $1766$ & $0.002$ \\ 
total\_precip\_std & bchave\_log & $0.210$ & $0.170$ & $0.260$ & $1766$ & $0$ \\ 
precip\_seasonality\_log\_std & mean\_temp\_std & $$-$0.170$ & $$-$0.210$ & $$-$0.120$ & $1766$ & $0$ \\ 
precip\_seasonality\_log\_std & temp\_seasonality\_log\_std & $0.070$ & $0.020$ & $0.120$ & $1766$ & $0.003$ \\ 
precip\_seasonality\_log\_std & sp\_rich\_raref\_log\_std & $$-$0.010$ & $$-$0.060$ & $0.040$ & $1766$ & $0.708$ \\ 
precip\_seasonality\_log\_std & shannon\_equit\_log\_std & $$-$0.240$ & $$-$0.280$ & $$-$0.200$ & $1766$ & $0$ \\ 
precip\_seasonality\_log\_std & cov\_height\_std & $$-$0.010$ & $$-$0.060$ & $0.050$ & $1326$ & $0.768$ \\ 
precip\_seasonality\_log\_std & cov\_dbh\_std & $$-$0.340$ & $$-$0.380$ & $$-$0.300$ & $1766$ & $0$ \\ 
precip\_seasonality\_log\_std & stems\_ha\_log\_std & $$-$0.260$ & $$-$0.300$ & $$-$0.210$ & $1766$ & $0$ \\ 
precip\_seasonality\_log\_std & bchave\_log & $$-$0.100$ & $$-$0.140$ & $$-$0.050$ & $1766$ & $0$ \\ 
mean\_temp\_std & temp\_seasonality\_log\_std & $$-$0.310$ & $$-$0.350$ & $$-$0.270$ & $1766$ & $0$ \\ 
mean\_temp\_std & sp\_rich\_raref\_log\_std & $$-$0.180$ & $$-$0.220$ & $$-$0.130$ & $1766$ & $0$ \\ 
mean\_temp\_std & shannon\_equit\_log\_std & $0$ & $$-$0.050$ & $0.040$ & $1766$ & $0.901$ \\ 
mean\_temp\_std & cov\_height\_std & $0.100$ & $0.050$ & $0.160$ & $1326$ & $0$ \\ 
mean\_temp\_std & cov\_dbh\_std & $0.140$ & $0.100$ & $0.190$ & $1766$ & $0$ \\ 
mean\_temp\_std & stems\_ha\_log\_std & $0.050$ & $0.010$ & $0.100$ & $1766$ & $0.024$ \\ 
mean\_temp\_std & bchave\_log & $$-$0.080$ & $$-$0.130$ & $$-$0.030$ & $1766$ & $0.001$ \\ 
temp\_seasonality\_log\_std & sp\_rich\_raref\_log\_std & $$-$0.340$ & $$-$0.380$ & $$-$0.300$ & $1766$ & $0$ \\ 
temp\_seasonality\_log\_std & shannon\_equit\_log\_std & $$-$0.130$ & $$-$0.180$ & $$-$0.090$ & $1766$ & $0$ \\ 
temp\_seasonality\_log\_std & cov\_height\_std & $$-$0.220$ & $$-$0.280$ & $$-$0.170$ & $1326$ & $0$ \\ 
temp\_seasonality\_log\_std & cov\_dbh\_std & $$-$0.140$ & $$-$0.190$ & $$-$0.100$ & $1766$ & $0$ \\ 
temp\_seasonality\_log\_std & stems\_ha\_log\_std & $$-$0.150$ & $$-$0.200$ & $$-$0.110$ & $1766$ & $0$ \\ 
temp\_seasonality\_log\_std & bchave\_log & $$-$0.140$ & $$-$0.190$ & $$-$0.100$ & $1766$ & $0$ \\ 
sp\_rich\_raref\_log\_std & shannon\_equit\_log\_std & $0.600$ & $0.570$ & $0.630$ & $1766$ & $0$ \\ 
sp\_rich\_raref\_log\_std & cov\_height\_std & $0.280$ & $0.230$ & $0.330$ & $1326$ & $0$ \\ 
sp\_rich\_raref\_log\_std & cov\_dbh\_std & $0.320$ & $0.270$ & $0.360$ & $1766$ & $0$ \\ 
sp\_rich\_raref\_log\_std & stems\_ha\_log\_std & $0.510$ & $0.480$ & $0.540$ & $1766$ & $0$ \\ 
sp\_rich\_raref\_log\_std & bchave\_log & $0.490$ & $0.460$ & $0.530$ & $1766$ & $0$ \\ 
shannon\_equit\_log\_std & cov\_height\_std & $0.100$ & $0.040$ & $0.150$ & $1326$ & $0$ \\ 
shannon\_equit\_log\_std & cov\_dbh\_std & $0.250$ & $0.200$ & $0.290$ & $1766$ & $0$ \\ 
shannon\_equit\_log\_std & stems\_ha\_log\_std & $0.610$ & $0.580$ & $0.640$ & $1766$ & $0$ \\ 
shannon\_equit\_log\_std & bchave\_log & $0.430$ & $0.390$ & $0.470$ & $1766$ & $0$ \\ 
cov\_height\_std & cov\_dbh\_std & $0.580$ & $0.550$ & $0.620$ & $1326$ & $0$ \\ 
cov\_height\_std & stems\_ha\_log\_std & $0.170$ & $0.110$ & $0.220$ & $1326$ & $0$ \\ 
cov\_height\_std & bchave\_log & $0.290$ & $0.240$ & $0.340$ & $1326$ & $0$ \\ 
cov\_dbh\_std & stems\_ha\_log\_std & $0.330$ & $0.280$ & $0.370$ & $1766$ & $0$ \\ 
cov\_dbh\_std & bchave\_log & $0.450$ & $0.410$ & $0.480$ & $1766$ & $0$ \\ 
stems\_ha\_log\_std & bchave\_log & $0.560$ & $0.530$ & $0.590$ & $1766$ & $0$ \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}