### 2019-11-21 analysis/output/include/corr_ci_tab.tex
Draft sent to supervisor

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Thu, Nov 21, 2019 - 14:28:02
\begin{table}[!htbp] \centering 
  \caption{} 
  \label{corr_ci_tab} 
\begin{tabular}{@{\extracolsep{5pt}} ccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
x\_var & y\_var & raw.r & raw.lower & raw.upper & n & p \\ 
\hline \\[-1.8ex] 
Sand \% & Org. C \% & $$-$0.500$ & $$-$0.540$ & $$-$0.470$ & $1766$ & p \textless 0.01 \\ 
Sand \% & CEC & $$-$0.560$ & $$-$0.600$ & $$-$0.530$ & $1766$ & p \textless 0.01 \\ 
Sand \% & MAP & $$-$0.490$ & $$-$0.520$ & $$-$0.450$ & $1766$ & p \textless 0.01 \\ 
Sand \% & MAP CoV & $0.330$ & $0.280$ & $0.370$ & $1766$ & p \textless 0.01 \\ 
Sand \% & MAT & $0.290$ & $0.250$ & $0.330$ & $1766$ & p \textless 0.01 \\ 
Sand \% & MAT CoV & $0.220$ & $0.180$ & $0.270$ & $1766$ & p \textless 0.01 \\ 
Sand \% & Sp. rich. & $$-$0.360$ & $$-$0.400$ & $$-$0.310$ & $1766$ & p \textless 0.01 \\ 
Sand \% & Shannon equit. & $$-$0.300$ & $$-$0.340$ & $$-$0.260$ & $1766$ & p \textless 0.01 \\ 
Sand \% & Tree height CoV & $$-$0.120$ & $$-$0.170$ & $$-$0.060$ & $1326$ & p \textless 0.01 \\ 
Sand \% & DBH CoV & $$-$0.190$ & $$-$0.230$ & $$-$0.140$ & $1766$ & p \textless 0.01 \\ 
Sand \% & Stems ha & $$-$0.270$ & $$-$0.310$ & $$-$0.230$ & $1766$ & p \textless 0.01 \\ 
Sand \% & AGB & $$-$0.230$ & $$-$0.280$ & $$-$0.190$ & $1766$ & p \textless 0.01 \\ 
Org. C \% & CEC & $0.300$ & $0.250$ & $0.340$ & $1766$ & p \textless 0.01 \\ 
Org. C \% & MAP & $0.440$ & $0.400$ & $0.480$ & $1766$ & p \textless 0.01 \\ 
Org. C \% & MAP CoV & $$-$0.310$ & $$-$0.350$ & $$-$0.270$ & $1766$ & p \textless 0.01 \\ 
Org. C \% & MAT & $$-$0.310$ & $$-$0.350$ & $$-$0.270$ & $1766$ & p \textless 0.01 \\ 
Org. C \% & MAT CoV & $$-$0.140$ & $$-$0.190$ & $$-$0.100$ & $1766$ & p \textless 0.01 \\ 
Org. C \% & Sp. rich. & $0.260$ & $0.210$ & $0.300$ & $1766$ & p \textless 0.01 \\ 
Org. C \% & Shannon equit. & $0.220$ & $0.180$ & $0.270$ & $1766$ & p \textless 0.01 \\ 
Org. C \% & Tree height CoV & $0.070$ & $0.020$ & $0.130$ & $1326$ & p \textless 0.01 \\ 
Org. C \% & DBH CoV & $0.140$ & $0.090$ & $0.190$ & $1766$ & p \textless 0.01 \\ 
Org. C \% & Stems ha & $0.170$ & $0.120$ & $0.210$ & $1766$ & p \textless 0.01 \\ 
Org. C \% & AGB & $0.250$ & $0.200$ & $0.290$ & $1766$ & p \textless 0.01 \\ 
CEC & MAP & $$-$0.010$ & $$-$0.060$ & $0.040$ & $1766$ & p = 0.66 \\ 
CEC & MAP CoV & $$-$0.510$ & $$-$0.540$ & $$-$0.480$ & $1766$ & p \textless 0.01 \\ 
CEC & MAT & $0.190$ & $0.150$ & $0.240$ & $1766$ & p \textless 0.01 \\ 
CEC & MAT CoV & $0.030$ & $$-$0.020$ & $0.070$ & $1766$ & p = 0.26 \\ 
CEC & Sp. rich. & $$-$0.020$ & $$-$0.060$ & $0.030$ & $1766$ & p = 0.46 \\ 
CEC & Shannon equit. & $0.220$ & $0.170$ & $0.260$ & $1766$ & p \textless 0.01 \\ 
CEC & Tree height CoV & $0.040$ & $$-$0.010$ & $0.090$ & $1326$ & p = 0.15 \\ 
CEC & DBH CoV & $0.240$ & $0.190$ & $0.280$ & $1766$ & p \textless 0.01 \\ 
CEC & Stems ha & $0.200$ & $0.150$ & $0.240$ & $1766$ & p \textless 0.01 \\ 
CEC & AGB & $0.030$ & $$-$0.020$ & $0.080$ & $1766$ & p = 0.23 \\ 
MAP & MAP CoV & $$-$0.120$ & $$-$0.160$ & $$-$0.070$ & $1766$ & p \textless 0.01 \\ 
MAP & MAT & $$-$0.120$ & $$-$0.170$ & $$-$0.070$ & $1766$ & p \textless 0.01 \\ 
MAP & MAT CoV & $$-$0.690$ & $$-$0.720$ & $$-$0.670$ & $1766$ & p \textless 0.01 \\ 
MAP & Sp. rich. & $0.430$ & $0.390$ & $0.470$ & $1766$ & p \textless 0.01 \\ 
MAP & Shannon equit. & $0.100$ & $0.060$ & $0.150$ & $1766$ & p \textless 0.01 \\ 
MAP & Tree height CoV & $0.250$ & $0.200$ & $0.300$ & $1326$ & p \textless 0.01 \\ 
MAP & DBH CoV & $0.150$ & $0.100$ & $0.190$ & $1766$ & p \textless 0.01 \\ 
MAP & Stems ha & $0.070$ & $0.030$ & $0.120$ & $1766$ & p \textless 0.01 \\ 
MAP & AGB & $0.210$ & $0.170$ & $0.260$ & $1766$ & p \textless 0.01 \\ 
MAP CoV & MAT & $$-$0.170$ & $$-$0.210$ & $$-$0.120$ & $1766$ & p \textless 0.01 \\ 
MAP CoV & MAT CoV & $0.070$ & $0.020$ & $0.120$ & $1766$ & p \textless 0.01 \\ 
MAP CoV & Sp. rich. & $$-$0.010$ & $$-$0.060$ & $0.040$ & $1766$ & p = 0.71 \\ 
MAP CoV & Shannon equit. & $$-$0.240$ & $$-$0.280$ & $$-$0.200$ & $1766$ & p \textless 0.01 \\ 
MAP CoV & Tree height CoV & $$-$0.010$ & $$-$0.060$ & $0.050$ & $1326$ & p = 0.77 \\ 
MAP CoV & DBH CoV & $$-$0.340$ & $$-$0.380$ & $$-$0.300$ & $1766$ & p \textless 0.01 \\ 
MAP CoV & Stems ha & $$-$0.260$ & $$-$0.300$ & $$-$0.210$ & $1766$ & p \textless 0.01 \\ 
MAP CoV & AGB & $$-$0.100$ & $$-$0.140$ & $$-$0.050$ & $1766$ & p \textless 0.01 \\ 
MAT & MAT CoV & $$-$0.310$ & $$-$0.350$ & $$-$0.270$ & $1766$ & p \textless 0.01 \\ 
MAT & Sp. rich. & $$-$0.180$ & $$-$0.220$ & $$-$0.130$ & $1766$ & p \textless 0.01 \\ 
MAT & Shannon equit. & $0$ & $$-$0.050$ & $0.040$ & $1766$ & p = 0.9 \\ 
MAT & Tree height CoV & $0.100$ & $0.050$ & $0.160$ & $1326$ & p \textless 0.01 \\ 
MAT & DBH CoV & $0.140$ & $0.100$ & $0.190$ & $1766$ & p \textless 0.01 \\ 
MAT & Stems ha & $0.050$ & $0.010$ & $0.100$ & $1766$ & p \textless 0.05 \\ 
MAT & AGB & $$-$0.080$ & $$-$0.130$ & $$-$0.030$ & $1766$ & p \textless 0.01 \\ 
MAT CoV & Sp. rich. & $$-$0.340$ & $$-$0.380$ & $$-$0.300$ & $1766$ & p \textless 0.01 \\ 
MAT CoV & Shannon equit. & $$-$0.130$ & $$-$0.180$ & $$-$0.090$ & $1766$ & p \textless 0.01 \\ 
MAT CoV & Tree height CoV & $$-$0.220$ & $$-$0.280$ & $$-$0.170$ & $1326$ & p \textless 0.01 \\ 
MAT CoV & DBH CoV & $$-$0.140$ & $$-$0.190$ & $$-$0.100$ & $1766$ & p \textless 0.01 \\ 
MAT CoV & Stems ha & $$-$0.150$ & $$-$0.200$ & $$-$0.110$ & $1766$ & p \textless 0.01 \\ 
MAT CoV & AGB & $$-$0.140$ & $$-$0.190$ & $$-$0.100$ & $1766$ & p \textless 0.01 \\ 
Sp. rich. & Shannon equit. & $0.600$ & $0.570$ & $0.630$ & $1766$ & p \textless 0.01 \\ 
Sp. rich. & Tree height CoV & $0.280$ & $0.230$ & $0.330$ & $1326$ & p \textless 0.01 \\ 
Sp. rich. & DBH CoV & $0.320$ & $0.270$ & $0.360$ & $1766$ & p \textless 0.01 \\ 
Sp. rich. & Stems ha & $0.510$ & $0.480$ & $0.540$ & $1766$ & p \textless 0.01 \\ 
Sp. rich. & AGB & $0.490$ & $0.460$ & $0.530$ & $1766$ & p \textless 0.01 \\ 
Shannon equit. & Tree height CoV & $0.100$ & $0.040$ & $0.150$ & $1326$ & p \textless 0.01 \\ 
Shannon equit. & DBH CoV & $0.250$ & $0.200$ & $0.290$ & $1766$ & p \textless 0.01 \\ 
Shannon equit. & Stems ha & $0.610$ & $0.580$ & $0.640$ & $1766$ & p \textless 0.01 \\ 
Shannon equit. & AGB & $0.430$ & $0.390$ & $0.470$ & $1766$ & p \textless 0.01 \\ 
Tree height CoV & DBH CoV & $0.580$ & $0.550$ & $0.620$ & $1326$ & p \textless 0.01 \\ 
Tree height CoV & Stems ha & $0.170$ & $0.110$ & $0.220$ & $1326$ & p \textless 0.01 \\ 
Tree height CoV & AGB & $0.290$ & $0.240$ & $0.340$ & $1326$ & p \textless 0.01 \\ 
DBH CoV & Stems ha & $0.330$ & $0.280$ & $0.370$ & $1766$ & p \textless 0.01 \\ 
DBH CoV & AGB & $0.450$ & $0.410$ & $0.480$ & $1766$ & p \textless 0.01 \\ 
Stems ha & AGB & $0.560$ & $0.530$ & $0.590$ & $1766$ & p \textless 0.01 \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}