### 2019-12-11 analysis/output/include/clust_summ.tex
Backup before aggressively refactoring code in light of Kyle's comments

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Wed, Nov 27, 2019 - 09:59:50
\begin{table}[!htbp] \centering 
  \caption{} 
  \label{clust_summ} 
\begin{tabular}{@{\extracolsep{5pt}} ccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
clust4 & c\_dom & c\_ind & n\_plots & sp\_rich\_raref & n\_stems\_ha & bchave \\ 
\hline \\[-1.8ex] 
1 & Julbernadia spp., Brachystegia spiciformis, Baikeaea plurijuga & Diplorhynchus condylocarpon, Burkea africana, Pseudolachnostylis maprouneifolia & 734 & 6(5) & 276(208) & 35.4(31.41) \\ 
2 & Julbernadia spp., Brachystegia spp., Isoberlinia angolensis & Julbernardia paniculata, Isoberlinia angolensis, Brachystegia longifolia & 728 & 8(5) & 272(195) & 47.6(44.01) \\ 
3 & Spirostachys africana, Senegalia spp., Euclea racemosa & Baikiaea plurijuga, Senegalia ataxacantha, Combretum collinum & 299 & 9(9) & 422(450.9) & 45(45.92) \\ 
4 & Colophospermum mopane & Colophospermum mopane, Combretum spp. & 102 & 4(4) & 322(228.8) & 43.8(36.49) \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}