### 2019-12-12 data_clean_flow/data_clean_flow.tex
Refactored most of the code

\begin{tikzpicture}[auto,scale=2, 
	round/.style={rounded rectangle,draw,thick,inner sep=5pt,minimum size=1cm, align=left},
	box/.style={rectangle,draw,thick,inner sep=5pt,minimum size=1cm, align=left},
	diam/.style={diamond,draw,thick,inner sep=5pt,minimum size=1cm, align=left, aspect=2},
	path/.style={->, thick, >=stealth'}]

\tikzset{mystyle/.style={->,double=black}}
\node  [round] (a)                    {Exclude plots with:\\- Charcoal and timber  harvesting\\- Farming/grazing\\- Fire manipulation};
\node  [box] (b) [right = 1.2cm of a] {Aggregate values of Zambian\\forestry commission plots};
\node  [round] (c) [below = 1.2cm of b] {Exclude plots where minimum\\measured tree size \textgreater{}5 cm};
\node  [round] (d) [left = 1.2cm of c] {Exclude stems:\\- \textless{}5 cm DBH\\- Dead\\- Unknown species};
\node  [box] (e) [below = 1.2cm of d] {Calculate mean plot\\level woodland traits};
\node  [round] (f) [right = 1.2cm of e] {Exclude plots with no\\tree species diversity};
\node  [round] (g) [below = 1.2cm of f] {Exclude compositional\\outlier plots};
\node  [box] (h) [left = 1.2cm of g] {Calculate reversed\\variables for SEM};
\node  [box] (i) [below = 1.2cm of h] {Transform variables to\\achieve normality};
\node  [box] (j) [right = 1.2cm of i] {Standardize variables};
\node  [diam] (k) [below = 1.2cm of j] {Analysis};
\draw [path] (a) -> (b);
\draw [path] (b) -> (c);
\draw [path] (c) -> (d);
\draw [path] (d) -> (e);
\draw [path] (e) -> (f);
\draw [path] (f) -> (g);
\draw [path] (g) -> (h);
\draw [path] (h) -> (i);
\draw [path] (i) -> (j);
\draw [path] (j) -> (k);

\end{tikzpicture}