### 2019-12-16 include/clust_summ.tex
Updated manuscript

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Mon, Dec 16, 2019 - 11:33:55
\begin{table}[!htbp] \centering 
  \caption{} 
  \label{clust_summ} 
\begin{tabular}{@{\extracolsep{5pt}} ccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
clust4 & c\_dom & c\_ind & n\_plots & n\_species\_raref & stems\_ha & agb\_ha \\ 
\hline \\[-1.8ex] 
1 & Julbernadia spp., Brachystegia spiciformis, Baikeaea plurijuga & Diplorhynchus condylocarpon, Burkea africana, Pseudolachnostylis maprouneifolia & 687 & 11(11.2) & 170(145.3) & 37(34.68) \\ 
2 & Julbernadia spp., Brachystegia spp., Isoberlinia angolensis & Julbernardia paniculata, Isoberlinia angolensis, Brachystegia longifolia & 757 & 18(17.6) & 215(171.7) & 48.8(43.7) \\ 
3 & Spirostachys africana, Senegalia spp., Euclea racemosa & Baikiaea plurijuga, Senegalia ataxacantha, Combretum collinum & 226 & 10(10) & 165(157.5) & 46(47.81) \\ 
4 & Colophospermum mopane & Colophospermum mopane, Combretum spp. & 99 & 7(8.2) & 190(155.7) & 41.5(36.93) \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}