### 2019-12-16 sem_path_diag/path_diagram_struc.tex

\begin{tikzpicture}[auto,scale=2, 
	observed/.style={rectangle,draw,thick,inner sep=5pt, outer sep=2pt, minimum size=1cm, align=center},
	latent/.style={circle, draw, thick, inner sep=0pt, outer sep=2pt, minimum size=2cm, align=center}, 
	path/.style={->, thick, >=stealth'},
	corr/.style={<->, thick, bend right=15, dotted, >=stealth'},
	loading/.style={->, thick, dotted, >=stealth'},
	loadingset/.style={->, thick, dashed, >=stealth'}]

\tikzset{mystyle/.style={->,double=black}}
\node [latent] (d) {Species\\Diversity}; 
\node [latent] (h) [right = 1in of d] {Struc.\\Diversity}; 
\node [observed] (b) [below = 1in of h] {AGB}; 
\node [observed] (i) [below = 1in of d] {Stem\\Density};

\coordinate[above= 0.7in of d] (domid);
\coordinate[above= 0.7in of h] (homid);
\node [observed] (ds) [left = 0.2in of domid] {Sp.\\Rich.};
\node [observed] (de) [right = 0.2in of domid] {Shannon\\Equit.};
\node [observed] (hd) [left = 0.2in of homid] {DBH\\CoV};
\node [observed] (hh) [right = 0.2in of homid] {Height\\CoV};

\draw [loadingset] (d) to node[left, pos=0.5] {\pcsdsd{}} (ds);
\draw [loading] (d) to node[right, pos=0.5] {\pcsded{}} (de);
\draw [loadingset] (h) to node[left, pos=0.5]  {\pcshdh{}} (hd);
\draw [loading] (h) to node[right, pos=0.5] {\pcshhh{}} (hh);

\draw [path] (d) to node {\pcsdh{}} (h);
\draw [path] (h) to node {\pcshb{}} (b);
\draw [corr] (d) to node {\pcsdi{}} (i);
\draw [path] (i) to node {\pcsib{}} (b);
\draw [path] (d) to node {\pcsdb{}} (b);

\end{tikzpicture}