### 2020-01-17 include/clust_summ.tex
Going through Kyle's comments

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Fri, Jan 17, 2020 - 12:26:04
\begin{table}[!htbp] \centering 
  \caption{} 
  \label{clust_summ} 
\begin{tabular}{@{\extracolsep{5pt}} ccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
clust4 & c\_dom & c\_ind & n\_plots & n\_species\_raref & stems\_ha & agb\_ha \\ 
\hline \\[-1.8ex] 
Marginal miombo & Julbernadia spp., Brachystegia spiciformis, Baikeaea plurijuga & Diplorhynchus condylocarpon, Burkea africana, Pseudolachnostylis maprouneifolia & 688 & 11(11.2) & 152(135.2) & 32.9(30.57) \\ 
Core miombo & Julbernadia spp., Brachystegia spp., Isoberlinia angolensis & Julbernardia paniculata, Isoberlinia angolensis, Brachystegia longifolia & 754 & 18(17.5) & 193(174.1) & 44.8(41.43) \\ 
Baikiaea & Spirostachys africana, Senegalia spp., Euclea racemosa & Baikiaea plurijuga, Senegalia ataxacantha, Combretum collinum & 226 & 10(10) & 162(152) & 45.3(47.36) \\ 
Mopane & Colophospermum mopane & Colophospermum mopane, Combretum spp. & 99 & 7(8.2) & 190(155.7) & 41.5(36.93) \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}