### 2020-01-17 manuscript/include/struc_model_fit_clust_stats.tex

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Fri, Jan 17, 2020 - 12:26:16
\begin{table}[!htbp] \centering 
  \caption{Model fit statistics for SEMs investigating the effects of tree diversity and stem density on AGB (\autoref{struc_mod}).} 
  \label{struc_model_fit_clust_stats} 
\begin{tabular}{@{\extracolsep{0pt}} ccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
{Cluster} & {n} & {$\chi^{2}$} & {DoF} & {CFI} & {TLI} & {LogLik} & {RMSEA} & {$R^{2}$ AGB} \\
\hline \\[-1.8ex] 
Marginal miombo & 525 & 44.750 & 6 & 0.966 & 0.916 & -3714.000 & 0.110 & 0.710 \\ 
Core miombo & 668 & 57.210 & 6 & 0.962 & 0.904 & -4224.000 & 0.100 & 0.680 \\ 
Baikiaea & 47 & 5.860 & 6 & 0.998 & 0.994 & -324.600 & 0.030 & 0.720 \\ 
Mopane & 84 & 9.420 & 6 & 0.971 & 0.927 & -591.600 & 0.080 & 0.450 \\ 
All & 1324 & 78.430 & 6 & 0.975 & 0.936 & -9119.000 & 0.090 & 0.690 \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}