### 2020-04-07 cover_letter/cover_letter.tex
Added some comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Thin Formal Letter
% LaTeX Template
% Version 1.11 (8/12/12)
% Compile with XeLaTeX

%----------------------------------------------------------------------------------------
%	DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[a4paper]{letter}

\usepackage[margin=2.5cm]{geometry}  %set margin width all around page

\addtolength{\voffset}{0in} % To push the letter further up on the page

\longindentation=0pt % Un-commenting this line will push the closing "Sincerely," to the left of the page

%----------------------------------------------------------------------------------------
%	 NAME & ADDRESS
%----------------------------------------------------------------------------------------

\signature{John L. Godlee} % Your name for the signature at the bottom

\address{Crew Building \\ Alexander Crum Brown Road \\ The King's Buildings \\ Edinburgh \\ EH9 3FF \\ United Kingdom \\ \\ johngodlee@gmail.com} % Your address and phone number

%----------------------------------------------------------------------------------------

\begin{document}

% Cover letter up to 250 words why the paper is of interest to readers of the journal

%----------------------------------------------------------------------------------------
%	ADDRESSEE 
%----------------------------------------------------------------------------------------

\begin{letter}{} % Name/title of the addressee

%----------------------------------------------------------------------------------------
%	LETTER CONTENT 
%----------------------------------------------------------------------------------------

	\opening{\textbf{Dear ,}}

I enclose a manuscript for submission to

We thank you for your consideration.

\vspace{\parskip} % Extra whitespace for aesthetics
\closing{Yours,}
\vspace{2\parskip} % Extra whitespace for aesthetics

% \ps{P.S. You can find additional information attached to this letter.} % Postscript text, comment this line to remove it

% \encl{Copyright permission form} % Enclosures with the letter, comment this line to remove it

%----------------------------------------------------------------------------------------

\end{letter}

\end{document}