### 2021-01-25 manuscript/title_page/title_page.tex

\documentclass[11pt,a4paper]{article}

% Define page geometry
\usepackage{geometry}
\geometry{left=2.2cm,
	right=2.2cm,
	top=2.2cm,
	bottom=2cm}
\parskip 0.15cm
\setlength{\parindent}{0cm}
\usepackage{pdflscape}
\usepackage[document]{ragged2e}

% Set font
\usepackage[T1]{fontenc}

% Image handling
\usepackage{graphics}  % Insert images easily
\usepackage{graphicx}  % Extended image support

\makeatletter
	\g@addto@macro\@floatboxreset\centering  % Automatically centre images (floats)
\makeatother

\graphicspath{ {img/} }
\usepackage{float}  %  Graphics placement [H] [H!] arguments
\usepackage{subfig}  % Compound figures

% Tables
\usepackage{multirow}
\usepackage{longtable}

% Bibliography management
\usepackage[UKenglish]{babel}
\usepackage[natbibapa]{apacite}
\bibliographystyle{apacite}

% Text formatting
\usepackage{url} % Allow nice formatting of URLs in text

\usepackage{enumerate}  % Enumerated lists

%\usepackage{lineno}  % Line numbers

\usepackage{textcomp}
\newcommand{\textapprox}{\raisebox{0.5ex}{\texttildelow}}  % Command for a good tilde

\usepackage{siunitx}
\usepackage{amsmath}

\usepackage{xcolor}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{#1}}}   % \todo{NOTE TO SELF WRITTEN IN RED}

\input{code_format}

% Custom title formatting
\let\oldtitle\title

\renewcommand{\title}[1]{\oldtitle{\vspace{-1.5cm}#1}}

\usepackage[breaklinks]{hyperref}
\definecolor{links}{RGB}{191,59,72}
\hypersetup{
	breaklinks,
	colorlinks,
	allcolors=links,
	linktoc=section,
	pdfauthor={John L. Godlee}
}
\def\subsectionautorefname{section}
\def\subsubsectionautorefname{section}
% \usepackage{biblatex}

\usepackage{dcolumn}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{calc,arrows,positioning,shapes,shapes.gates.logic.US,trees}
\usepackage{pdflscape}
\usepackage{longtable}
\usepackage{fmtcount}

\begin{document}

{\textbf{Structural diversity and tree density drives variation in the biodiversity-ecosystem function relationship of woodlands and savannas}}

\vspace{1cm}

John L. Godlee\textsuperscript{1}, Casey M. Ryan\textsuperscript{1}, David Bauman\textsuperscript{2,3}, Samuel J. Bowers\textsuperscript{1}, Joao M. B. Carreiras\textsuperscript{4}, Antonio Valter Chisingui\textsuperscript{5}, Joris P. G. M. Cromsigt\textsuperscript{6,7,8}, Dave J. Druce\textsuperscript{9,10}, Manfred Finkch\textsuperscript{11}, Francisco Maiato Gon\c{c}alves\textsuperscript{5}, Ricardo M. Holdo\textsuperscript{12}, Steve Makungwa\textsuperscript{13}, Iain M. McNicol\textsuperscript{1}, Edward T. A. Mitchard\textsuperscript{1}, Anderson Muchawona\textsuperscript{14}, Rasmus Revermann\textsuperscript{11,15}, Natasha Sofia Ribeiro\textsuperscript{16}, Abel Siampale\textsuperscript{17}, Stephen Syampungani\textsuperscript{18}, Jos\'{e} Jo\~{a}o Tchamba\textsuperscript{5}, Hemant G. Tripathi\textsuperscript{19}, Johannes Wallenfang\textsuperscript{11}, Mariska te Beest\textsuperscript{20,21,22}, Mathew Williams\textsuperscript{1}, Kyle G. Dexter\textsuperscript{1,23}

1: School of GeoSciences, University of Edinburgh, Edinburgh, EH9 3FF, United Kingdom \\
2: Environmental Change Institute, School of Geography and the Environment, University of Oxford, Oxford, United Kingdom \\
3: Laboratoire d'\'{E}cologie V\'{e}g\'{e}tale et Biogéochimie, CP 244, Universit\'{e} Libre de Bruxelles, Brussels, Belgium \\
4: National Centre for Earth Observation (NCEO), University of Sheffield, Hicks Building, Hounsfield Road, Sheffield, S3 7RH, United Kingdom \\
5: Herbarium of Lubango, ISCED Hu\'{i}la, Sarmento Rodrigues Str. No. 2, CP. 230, Lubango, Angola \\
6: Department of Wildlife, Fish, and Environmental Studies, Swedish University of Agricultural Sciences, Ume\aa, Sweden \\
7: Department of Zoology, Centre for African Conservation Ecology, Nelson Mandela University, Port Elizabeth, South Africa \\
8: Copernicus Institute of Sustainable Development, Utrecht University, 3584CS Utrecht, the Netherlands \\
9: Ecological Advice, Ezemvelo KZN Wildlife, Hluhluwe-iMfolozi Park, South Africa \\
10: School of Life Sciences, University of KwaZulu-Natal, Pietermaritzburg, South Africa \\
11: Biodiversity, Evolution and Ecology of Plants, Institute of Plant Sciences and Microbiology, University of Hamburg, Ohnhorststr. 18, 22609 Hamburg, Germany \\
12: Odum School of Ecology, University of Georgia, 140 E. Green St., Athens, GA 30602, USA \\
13: Lilongwe University of Agriculture and Natural Resources (LUANAR), Lilongwe, Malawi \\
14: Forest Research Centre, 1 Orange Groove Highlands, Harare, Zimbabwe \\
15: Faculty of Natural Resources and Spatial Sciences, Namibia University of Science and Technology, Windhoek, Namibia \\
16: Department of Forest Engineering, Faculty of Agronomy and Forest Engineering, Universidade Eduardo Mondlane, Av. Julius Nyerere, 3453, Campus Universitario, Maputo, Mozambique \\
17: Forestry Department Headquarters - Ministry of Lands and Natural Resources, Cairo Road, Lusaka, Zambia \\
18: School of Natural Resources, Copperbelt University, Kitwe, Zambia \\
19: Faculty of Biological Sciences, University of Leeds, LS2 9JT, United Kingdom\\
20: Copernicus Institute of Sustainable Development, Utrecht University, 3508 TC Utrecht, The Netherlands \\
21: Centre for African Conservation Ecology, Nelson Mandela University, Port Elizabeth, 6031, South Africa \\
22: South African Environmental Observation Network, Grasslands-Forests-Wetlands Node, Montrose, 3201, South Africa \\
23: Royal Botanic Garden Edinburgh, Edinburgh, EH3 5LR, United Kingdom \\

\newpage{}

\textbf{Author for correspondence:}

John L. Godlee \\
Telephone: +44(0) 7397 978 688 \\
Email: johngodlee@gmail.com \\

\vspace{1cm}

\textbf{Word count:}

\vspace{0.2cm}

\begin{tabular}{r|r}
Main body word count & 6384\\
Summary & 193\\
Introduction & 1490\\
Materials and Methods & 1999\\
Results & 726\\
Discussion & 2216\\
Acknowledgements & 122\\
\end{tabular}

\vspace{1cm}

Number of figures: 7

Number of tables: 2

Colour figures: 1, 3, 5, 6

Number of supporting information files: 5

\end{document}