### 2021-07-09 manuscript/title_page/title_page.tex

\documentclass[11pt,a4paper]{article}

% Define page geometry
\usepackage{geometry}
\geometry{left=2.2cm,
	right=2.2cm,
	top=2.2cm,
	bottom=2cm}
\parskip 0.2cm
\setlength{\parindent}{0cm}

% Set font
\usepackage[T1]{fontenc}

\begin{document}

{\textbf{Title: Structural diversity and tree density drives variation in the biodiversity-ecosystem function relationship of woodlands and savannas}}

John L. Godlee\textsuperscript{1}, Casey M. Ryan\textsuperscript{1}, David Bauman\textsuperscript{2,3}, Samuel J. Bowers\textsuperscript{1}, Joao M. B. Carreiras\textsuperscript{4}, Antonio Valter Chisingui\textsuperscript{5}, Joris P. G. M. Cromsigt\textsuperscript{6,7,8}, Dave J. Druce\textsuperscript{9,10}, Manfred Finckh\textsuperscript{11}, Francisco Maiato Gon\c{c}alves\textsuperscript{5}, Ricardo M. Holdo\textsuperscript{12}, Steve Makungwa\textsuperscript{13}, Iain M. McNicol\textsuperscript{1}, Edward T. A. Mitchard\textsuperscript{1}, Anderson Muchawona\textsuperscript{14}, Rasmus Revermann\textsuperscript{11,15}, Natasha Sofia Ribeiro\textsuperscript{16}, Abel Siampale\textsuperscript{17}, Stephen Syampungani\textsuperscript{18}, Jos\'{e} Jo\~{a}o Tchamba\textsuperscript{5}, Hemant G. Tripathi\textsuperscript{19}, Johannes Wallenfang\textsuperscript{11}, Mariska te Beest\textsuperscript{20,21,22}, Mathew Williams\textsuperscript{1}, Kyle G. Dexter\textsuperscript{1,23}

\section*{Affiliations}

1: School of GeoSciences, University of Edinburgh, Edinburgh, EH9 3FF, United Kingdom \\
2: Environmental Change Institute, School of Geography and the Environment, University of Oxford, Oxford, OX1 3QY, United Kingdom \\
3: Laboratoire d'\'{E}cologie V\'{e}g\'{e}tale et Biogéochimie, CP 244, Universit\'{e} Libre de Bruxelles, Brussels, B-1050, Belgium \\
4: National Centre for Earth Observation (NCEO), University of Sheffield, Hicks Building, Hounsfield Road, Sheffield, S3 7RH, United Kingdom \\
5: Herbarium of Lubango, ISCED Hu\'{i}la, Sarmento Rodrigues Str. No. 2, CP. 230, Lubango, Angola \\
6: Department of Wildlife, Fish, and Environmental Studies, Swedish University of Agricultural Sciences, Ume\aa, SE-907 36, Sweden \\
7: Department of Zoology, Centre for African Conservation Ecology, Nelson Mandela University, Port Elizabeth, 6031, South Africa \\
8: Copernicus Institute of Sustainable Development, Utrecht University, Utrecht, 3584CS, The Netherlands \\
9: Ecological Advice, Ezemvelo KZN Wildlife, Hluhluwe-iMfolozi Park, 3202, South Africa \\
10: School of Life Sciences, University of KwaZulu-Natal, Pietermaritzburg, 3209, South Africa \\
11: Biodiversity, Evolution and Ecology of Plants, Institute of Plant Sciences and Microbiology, University of Hamburg, Ohnhorststr. 18, Hamburg, 22609, Germany \\
12: Odum School of Ecology, University of Georgia, 140 E. Green St., Athens, GA 30602, USA \\
13: Lilongwe University of Agriculture and Natural Resources (LUANAR), Lilongwe, Malawi \\
14: Forest Research Centre, 1 Orange Groove Highlands, Harare, Zimbabwe \\
15: Faculty of Natural Resources and Spatial Sciences, Namibia University of Science and Technology, Windhoek, 10005, Namibia \\
16: Department of Forest Engineering, Faculty of Agronomy and Forest Engineering, Universidade Eduardo Mondlane, Av. Julius Nyerere, 3453, Campus Universitario, Maputo, Mozambique \\
17: Ministry of Lands and Natural Resources, Cairo Road, Lusaka, Zambia \\
18: School of Natural Resources, Copperbelt University, Kitwe, Zambia \\
19: Faculty of Biological Sciences, University of Leeds, LS2 9JT, United Kingdom\\
20: Copernicus Institute of Sustainable Development, Utrecht University, Utrecht, 3508, The Netherlands \\
21: Centre for African Conservation Ecology, Nelson Mandela University, Port Elizabeth, 6031, South Africa \\
22: South African Environmental Observation Network, Grasslands-Forests-Wetlands Node, Montrose, 3201, South Africa \\
23: Royal Botanic Garden Edinburgh, Edinburgh, EH3 5LR, United Kingdom \\

\newpage
\section*{ORCiD}

John L. Godlee - 0000-0001-5595-255X \\
Casey M. Ryan - 0000-0002-1802-0128 \\
David Bauman - 0000-0001-9115-6518 \\
Samuel J. Bowers - 0000-0001-5418-0359 \\
Joao M. B. Carreiras - 0000-0003-2737-9420 \\
Antonio Valter Chisingui - 0000-0002-5137-1448 \\
Joris P. G. M. Cromsigt - 0000-0002-8632-9469 \\
Dave J. Druce - 0000-0003-2976-6029 \\
Manfred Finckh - 0000-0003-2186-0854 \\
Francisco Maiato Gon\c{c}alves - 0000-0002-8859-7491 \\
Ricardo M. Holdo - 0000-0002-2484-0587 \\
Steve Makungwa -  \\
Iain M. McNicol - 0000-0003-3073-5668 \\
Edward T. A. Mitchard - 0000-0002-5690-4055 \\
Anderson Muchawona -  \\
Rasmus Revermann - 0000-0002-7044-768X \\
Natasha Sofia Ribeiro - 0000-0002-5369-5905 \\
Abel Siampale -  \\
Stephen Syampungani -  \\
Jos\'{e} Jo\~{a}o Tchamba - 0000-0002-3770-2482 \\
Hemant G. Tripathi - 0000-0001-6045-4336 \\
Johannes Wallenfang -  \\
Mariska te Beest - 0000-0003-3673-4105 \\
Mathew Williams - 0000-0001-6117-5208 \\
Kyle G. Dexter - 0000-0001-9232-5221 \\

\section*{Author for correspondence}

John L. Godlee \\
Telephone: +44(0) 7397 978 688 \\
Email: johngodlee@gmail.com \\

\section*{Word count}

Main body word count: 6706 \\
Summary: 198 \\
Introduction: 1107 \\
Materials and Methods: 2089 \\
Results: 756 \\
Discussion: 2226 \\
Conclusion: 330 \\
Acknowledgements: 122 \\
Number of figures: 7 \\
Number of tables: 2 \\
Colour figures: 1, 3, 5, 6 \\
Number of supporting information files: 5 \\

\end{document}